---
title: 'Problem Set #1: MVIS 5301'
author: "C. Todd Lombardo"
date: "11 January 2015"
output: html_document
---
This work can also be found online at: <http://rpubs.com/ctodd/PS1>  

***
###1. What is the level of measurement for each of these variables (from the 2002 General Social Survey) (nominal, ordinal, interval/ratio)? Briefly justify your answer. *Note: Ignore the response categories of NAP, DK, and NA for purposes of this question.*

#####CHILDS – How many children have you ever had? 
  0-6. Actual Number; 
  7. Seven or more; 
  9. DK or NA  
  
Response  |	TOTAL	|	%
-----------------	|	-----------------	|	-----------------
0) None	|	799	|	28.9
1) One	|	469	|	17
2) Two	|	657	|	23.8
3) Three	|	481	|	17.4
4) Four	|	185	|	6.7
5) Five	|	73	|	2.6
6) Six	|	40	|	1.4
7) Seven	|	22	|	0.8
8) Eight or more	|	34	|	1.2
9) No answer	|	5	|	0.2
TOTAL	|	2765	|	100

> *Ordinal.* At first glance, one might think the levels of measurement with these variables is *interval/ratio*, since a zero means an absence of children and there are equal distances between each measure. There are no "half-children", and if you have four children you have twice as many than if you only have two. However, the level of measurement is actually *ordinal* for this question since there is a grouping of responses at 'Eight or more' children. There's no difference if you have 8, 9 or 15 children for this grouping, but there is a difference between this grouping and any response less than eight, so it would not be considered *nominal* either.

#####EDUC – Highest year of school completed 
  0-20. Actual number of years; 
  9. DK or NA

Response  |	TOTAL	|	%
------------- | -------------| -------------
0) No formal schooling	|	5	|	0.2
1) 1st grade	|	2	|	0.1
2) 2nd grade	|	15	|	0.5
3) 3rd grade	|	2	|	0.1
4) 4th grade	|	6	|	0.2
5) 5th grade	|	8	|	0.3
6) 6th grade	|	25	|	0.9
7) 7th grade	|	14	|	0.5
8) 8th grade	|	66	|	2.4
9) 9th grade	|	60	|	2.2
10) 10th grade	|	88	|	3.2
11) 11th grade	|	137	|	5
12) 12th grade	|	818	|	29.6
13) 1 year of college	|	265	|	9.6
14) 2 years	|	374	|	13.5
15) 3 years	|	157	|	5.7
16) 4 years	|	377	|	13.6
17) 5 years	|	93	|	3.4
18) 6 years	|	128	|	4.6
19) 7 years	|	43	|	1.6
20) 8 years	|	70	|	2.5
98) Don't know	|	5	|	0.2
99) No answer	|	7	|	0.3
TOTAL	|	2765	|	100

> Levels of measurement here are *interval/ratio* as a zero level indicates no years of schooling and the intervals equal in length. Four years of schooling is twice as long as 2 years of schooling.
>
>


***
###2.  A student discovers that his grade on a recent test was in the 72nd percentile. If 90 students took the test, then approximately how many students received a higher grade than he did? 
> Approximately 25 students
```{r}
90-(90 * .72)
````
***
###3.  A sample of 7 underweight babies was fed a special diet and the following weight gains (lbs) were observed at the end of three months.

`6.7   2.7   2.5   3.6   3.4   4.1   4.8`

####Find the mean and standard deviation for these 7 babies. Show your calculations (credit will not be given if you simply provide the answers).   

> First, we find the mean (u), then we substract each variable from the mean (x-u), then square each

```{r}
x=c(6.7,2.7,2.5,3.6,3.4,4.1,4.8)       # Entering the Dataset into a Vector

u=sum(x)/length(x)                      # Calculate the Mean

u                                       # Print the Mean
````

x| x-u| (x-u)^2  
------------- | -------------| -------------
6.7|  2.7  |7.4
2.7  |-1.3|	1.6
2.5|	-1.5|	2.2
3.6|-0.4	|0.1
3.4|-0.6	|0.3
4.1|  0.1	|0.0
4.8|  0.8	|0.7

> We then sum those up, divide by number of variables minus one to calculate the variance, and take the square root to get the standard deviation.

```{r}
Variance = sum((x-u)^2)/(length(x)-1)    # Calculate Variance

sqrt(Variance)                          # Calculate and print Standard Deviation
````

***
###4.  The Nielsen Company publishes information on the TV-viewing habits of Americans in Nielsen Report on Television. A sample of 20 people yielded the weekly viewing times, in hours, in the table below. 

####a)  Determine the median of these data.
```{r}
set1=c(25,41,27,32,43,66,35,31,15,5,34,26,32,38,16,30,38,30,20,21) # Dataset into a Vector
n=length(set1)              # Obtain number of variables
u=sum(set1)/length(set1)    # Mean
sorted = sort(set1)         # Sort the data into a new vector
                            # The median is the average of the two middle values, 30 and 31
medianvalues = sorted[(length(set1)/2)] + (sorted[(length(set1)/2)]+1)
medianvalues / 2 
```
####b)  The quartiles of these data.
```{r}
# Q1 = 25th Percentile
quantile(sorted, 0.25)
# Q3 = 75th Percentile
quantile(sorted, 0.75)
# IQR = Q3 - Q1   <-- The interquartile range
IQR(sorted)
```
####c)	Obtain the lower and upper limits.
> Lower Limit
```{r}
24 - 1.5*(IQR(sorted))               # Lower Limit = Q1 – 1.5 × IQR
```
> Upper Limit
```{r}
35.75 + 1.5*(IQR(sorted))            # Upper Limit = Q3 + 1.5 × IQR
```

####d)	Determine potential outliers, if any.  

> Potential outliers are 5 and 66 since they are outside the range of 6.375 and 53.375. A boxplot can easily visualize this, and shows that 66 is indeed an outlier.  
```{r echo=FALSE}
boxplot(set1)
```

---
####5.	Use the Area under the standard normal curve table to obtain the shaded area under the standard normal curves below. 

####a. Area between z = -1.28 and 1.28

> The area under the curve for z = -1.28 is 0.1003. Multiply by two, then substract from 1:
> 1 - (2 *(0.1003)) = 0.7994 or 79.94% of the area under the normal distribution curve.

####b. Area between z = -1.64 and 1.64 

> The area under the curve for z =-1.64 is 0.0505. Similar to case a: multiply by two, then substract from 1:
> 1 - (2 *(0.0505)) = 0.899 or 89.90% of the area under the normal distribution curve.

####c. Area below z = -1.96 and above 1.96

> The area under the curve for z = -1.96 is 0.0250, so the area would be double that: 0.05, or 5% of the total area under the normal distribution curve.

####d. Area below z = -2.33 and 2.33

> The area under the curve for z = -2.33 is 0.0102, so the area would be double that: 0.0204, or 2.04% of the total area under the normal distribution curve.


[id]: "Problem Set #1 MVIS 5301 Statistical Applications for Visualization"
